### 2019-05-09 papers/issac19demo/main.tex
small changes to the software demo abstract
* Sage → SageMath
* official affiliation and funding sources for MM
* formatting changes
* explicitly refer to v0.4
* installation instructions
* reword a couple of sentences, mention some other improvements in 0.4,
  work that went directly into Sage
* smallsteps → ssw (cf. previous commit)
* thank reviewers of Sage patches

\documentclass[11pt]{article}
\usepackage[fleqn]{amsmath}
\usepackage{sigsam, sagetex, tikz}
\usepackage{hyperref}
\usepackage{booktabs}

% PAGE LIMIT: FOUR PAGES!

% leave as is
\issue{Vol.xx, No.xx, Issue xxx, month 201x}
\articlehead{ISSAC 2019 abstracts}
\titlehead{Multivariate Ore Polynomials in SageMath}
\authorhead{Manuel Kauers, Marc Mezzarobba}
%\titlehead{Title of your paper}
%\authorhead{Author's Name}
\setcounter{page}{1}

\def\<#1>{\langle#1\rangle}
\let\Bold\mathbb
\let\set\mathbb

% \setkeys{FV}{vspace=0pt}
\makeatletter
\let\FV@ListVSpace\relax
\makeatother

\date{}

\begin{document}

\title{Multivariate Ore Polynomials in SageMath}

\author{%
  Manuel Kauers\footnote{Supported by the Austrian FWF grants P31571-N32, F5004} \
  and Marc Mezzarobba%
  \footnote{Supported in part by ANR grant ANR-14-CE25-0018-01 (FastRelax).}%
  \\[\medskipamount]
  Institute for Algebra $\cdot$ Johannes Kepler University $\cdot$ Linz, Austria\\
  Sorbonne Université, CNRS, Laboratoire d'informatique de Paris~6, LIP6,
  F-75005 Paris, France\\[\medskipamount]
  \href{mailto:manuel@kauers.de}{\nolinkurl{manuel@kauers.de}} $\cdot$
  \href{mailto:marc@mezzarobba.net}{\nolinkurl{marc@mezzarobba.net}}
}

\maketitle

\begin{abstract}
  We present the latest update of the ore\_algebra package for SageMath.
  The main new feature in this release is the support of operators in several variables.
\end{abstract}

\section{Introduction}

Ore polynomials are important features of computer algebra systems because they provide
functionality for manipulating differential and recurrence operators that describe
so-called D-finite functions and sequences~\cite{stanley80,kauers10j,kauers13}.
Packages for working with Ore polynomials include, among others,
OreTools~\cite{abramov03} and gfun~\cite{salvy94} for Maple,
the HolonomicFunctions package of Koutschan~\cite{koutschan10c} for Mathematica,
and ore\_algebra package by Kauers, Jaroschek and Johansson~\cite{kauers14b} for
SageMath (Sage)~\cite{zimmermann18}. This software demo abstract is about the most recent version of the
latter package, which works with Sage~8.7.

Earlier versions of the ore\_algebra package already provided functionality for basic arithmetic
and actions; GCRD and LCLM; D-finite closure properties; natural transformations between related
algebras; guessing; desingularization; solvers for polynomials, rational functions and (generalized)
power series. Mezzarobba~\cite{mezzarobba16} added a subpackage ore\_algebra.analytic for symbolic-numeric
computations with D-finite analytic functions, which is now fully integrated into the main version of the package.
Hofstadler~\cite{hofstadler19} contributed a set of convenience functions which support a better integration
with symbolic expressions.

While basic arithmetic of Ore polynomials in several variables has been supported since the
beginning, everything else so far was limited to the univariate case.
The main new feature in versions~0.3--0.4 of the package is an extension of the core
functionality to multivariate operators, with efficient support of D-finite closure properties and
creative telescoping as well as a rudimentary implementation of Gr\"obner bases.
Other novelties include substantially improved numerical evaluation code compared to the status
described in~\cite{mezzarobba16}, and preliminary support for Python~3.

\section{Basic Examples}

With a typical Sage setup, ore\_algebra~v0.4 can be installed using the shell command
\begin{verbatim}
  $ sage -pip install git+https://github.com/mkauers/ore_algebra.git@0.4
\end{verbatim}
(Omit the \verb!@0.4! suffix for the latest development version.)

After loading the package and creating the base ring $\set Z[x,y]$, the Ore algebra of differential
operators for $x$ and $y$ can be created as follows. 
\begin{sageexample}
  sage: from ore_algebra import *
  sage: R.<x,y> = PolynomialRing(ZZ)
  sage: A.<Dx,Dy> = OreAlgebra(R)
  sage: A
\end{sageexample}
The usual arithmetic respects the required commutation rules. 
\begin{sageexample}
  sage: (Dx + Dy)*(y + x)
  sage: (x + y)*(Dx + Dy)
\end{sageexample}
Differential operators can be applied to objects which Sage knows to differentiate and to multiply
by base ring elements:
\begin{sageexample}
  sage: (Dx + Dy)(sin(x*y))
\end{sageexample}
Using $Sx,Sy$ instead of $Dx,Dy$ creates an Ore algebra with shift operators instead of differential operators.
Of course, combinations like $\set Z[x,y]\<Sx,Dy>$ and other generators besides shift and derivation are also possible.
The available operators are the same as in the univariate case, see~\cite{kauers14b} for more details.

In the univariate case, methods for executing closure properties were directly attached to the annihilating operators.
In the multivariate case, where several annihilating operators are needed to describe a D-finite object, we must work
with left ideals of Ore algebras. Ideals can be constructed as follows.
\begin{sageexample}
  sage: I = A.ideal([Dy^2, Dx*Dy, Dx^2 - y*Dy + 1])
  sage: I
\end{sageexample}
Note that the base ring is always converted to a field.

Various ideal theoretic operations are available as methods. 
\begin{sageexample}
  sage: I.eliminate([Dy])
  sage: Dx^3 + Dx in I # (membership)
  sage: J = A.ideal([-Dy^2 + x*Dx - 1, Dx*Dy, Dx^2])
  sage: I == J
  sage: K = I.intersection(J) # closure property `addition`
  sage: K <= J # (inclusion)
  sage: K = I.symmetric_product(I) # closure property `multiplication`
  sage: K.vector_space_basis() # terms under the staircase
  sage: I.annihilator_of_associate(Dx + Dy) # closure property `ore action`
  sage: I.annihilator_of_composition(x=x^2,y=1-x^2) # closure property `composition`
\end{sageexample}
Most of these operations rely on a generic FGLM implementation which uses the packages own linear system solvers.
There is also a direct implementation of Buchberger's algorithm for computing Gr\"obner bases, which however is
not highly optimized. 
\begin{sageexample}
  sage: I.groebner_basis()
\end{sageexample}

All operations discussed so far are generalizations of the univariate counterparts that already existed
in earlier versions. A feature that is inherently multivariate is creative telescoping~\cite{zeilberger91,chyzak00,chyzak14}.
Here, the input is a left ideal $I\subseteq A$, say $A=\set Q(x,y)\<Dx,Dy>$ and the task is to find
the ideal consisting of all operators $P\in\set Q(x)\<Dx>$ such that there exists $Q\in A$ with $P-D_yQ\in I$.
The operators $P$ are called \emph{telescopers,} and the corresponding operators $Q$ are called their \emph{certificates.}
Creative telescoping is a central operation for evaluating definite sums and integrals~\cite{chyzak14}.
We cannot explain here in detail the relevance of creative telescoping for these applications, but the use of the
package is easily demonstrated:
\begin{sageexample}
  sage: A.ideal([Dx - 2*x*y^2, Dy - 2*x^2*y]).ct(Dy)
\end{sageexample}
The output is a pair $([P_1,P_2,\dots],[Q_1,Q_2,\dots])$ whose first component is a basis of the ideal of telescopers
and whose second component contains the certificates corresponding to the basis elements.
The example above originates from the integration problem $\int_y\exp(x^2y^2)dy$. The following example is the essence
of a proof of the binomial theorem $\sum_k\binom nk=2^n$:
\begin{sageexample}
  sage: B.<Sx,Sy> = OreAlgebra(R)
  sage: B.ideal([(y+1)*Sy + (y-x), (x-y+1)*Sx - (x+1)]).ct(Sy - 1)
\end{sageexample}

Finally, also the guessing features have been extended to the multivariate setting. For example, annihilating operators
for the binomial coefficients can be found as follows.
\begin{sageexample}
  sage: data = [[binomial(n,k) for k in range(20)] for n in range(20)]
  sage: guess(data, B, point_filter = lambda n,k : k<=n).groebner_basis()
\end{sageexample}
The option \verb|point_filter| informs the guessing engine which part of the input array contains interesting data.
A number of further options is specified in the documentation.

\section{Some Slightly Larger Examples}

The examples shown above need virtually no computation time. In order to get some idea about the performance of our
implementation, we consider a set of 19 creative telescoping problems that appeared in a study of restricted lattice
walks~\cite{bostan16b}.
Each of the 19 problems starts from a certain rational function $r\in\set Q(u,v,t)$ and the task consists of two
applications of creative telescoping. Writing $I$ for the ideal of annihilating operators of $r$ in $\set Q(u,v,t)\<Du,Dv,Dt>$,
we first have to compute the ideal $T_v$ in $\set Q(u,t)\<Du,Dt>$ of all telescopers of $I$ with respect to~$Dv$ and then
the ideal $T_{u,v}$ in $\set Q(t)\<Dt>$ of telescopers of $T_v$ with respect to~$Du$. The smallest rational function in
the collection is
\[
\frac{{-}u^2 v^2{+}u^2{+}v^2{-}1}{t u^3 v^2{-}t u^3 v{+}t u^2 v^3{-}2 t u^2 v^2{+}2 t u^2 v{-}t u^2{-}t u v^3{+}2 t u v^2{-}2 t
  u v+t u-t v^2+t v-u^2 v^2+u^2 v+u v^2-u v},
\]
and the largest is about twice as long. The complete data is available in ore\_algebra.examples.ssw.

In the table below, we compare our performance to that of Koutschan's package~\cite{koutschan10c}. We do not claim that a
performance comparison will look similarly for any creative telescoping problem, but we think the experiment is at least
an indication that our implementation is not too bad.
Note in passing that reaching an acceptable performance level required a number of improvements to
the implementation of multivariate polynomials and rational functions in Sage itself.
Although ore\_algebra is developed as an external Sage package, a significant part of the
development effort reported here will thus directly benefit other Sage users.

\begin{center}
  % timings taken on picard.algebra.uni-linz.ac.at
  % with Sage 8.7 and ore_algebra version 7ab0f7e821cde38e1ace6ae0844bc34e726aaf5a
  % with Mathematica 12 and HolonomicFunctions.m version 1.7.3
  % on 2019-04-25
  \begin{tabular}{rccccccccccccccccccc}
    \toprule
    case & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 \\\midrule
    ore\_algebra & 601ms & 668ms & 705ms & 862ms & 4.82s & 10.3s & 5.22s & 10.7s & 11.1s & 11.9s \\
    HolonomicFunctions.m & 1.55s & 1.59s & 1.86s & 2.09s & 25.2s & 51.9s & 26.8s & 55.1s & 51.5s &
    63.4s \\
    \midrule
    case & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 \\\midrule
    ore\_algebra & 4.46s & 8.57s & 11.1s & 11.7s & 4.21s & 8.27s & 18.9s & 36.2s & 2.05s \\
    HolonomicFunctions.m & 19.4s & 35.4s & 51.3s & 62.3s & 19.1s & 34.7s & 198s & 445s & 5.77s \\
    \bottomrule
  \end{tabular}
\end{center}

\paragraph{Acknowledgements.}

We thank Bruno Grenet, Ben Hutz, Julian Rüth, and Travis Scrimshaw for reviewing the Sage patches
mentioned above.

\small
\bibliographystyle{plain}
\bibliography{bib}

\end{document}